---
output: md_document
---

```{r include = FALSE}
blog_dir <- strsplit(getwd(), "content")[[1]][1]
base_dir <- paste0(blog_dir, "static")
base_url <- "/"
fig_path <- "images/"

knitr::opts_knit$set(base.dir = base_dir, base.url = base_url)
knitr::opts_chunk$set(fig.path = fig_path)
```
# Inflation and Currency Conversions

## Introduction
We often need to adjust costs into specific currencies and years.

## When might I use this?
If needing to convert to unit or output costs into US dollars, local currency or international dollars using Gross Domestic Product values, in present or historical values. 

# Inflation and cost data packages

## Package: wbstats
Programmatic Access to Data and Statistics from the World Bank
API

### Maintained by
Jesse Piburn <piburnjo@ornl.gov>

## URLs
https://cran.r-project.org/web/packages/wbstats/wbstats.pdf
https://cran.r-project.org/web/packages/wbstats/vignettes/Using_the_wbstats_package.html

## What does this package do?
The package contains downloadable data from the World Bank.

## How do I input my data to it/what inputs does it take?
Not applicable

## What outputs do I get?
See https://cran.r-project.org/web/packages/wbstats/vignettes/Using_the_wbstats_package.html 

## Sample code section
```{r}
library(wbstats)
library(data.table)

pop.data <- wb(indicator = "NY.GDP.DEFL.KD.ZG")
pop.data <- as.data.table(pop.data)


pop.data <- pop.data[country=="World"]  
pop.data[ , date := as.numeric(date)]
pop.data[ , index := value]
pop.data[ , year := date]

# adapted from - inflatbly package - https://github.com/HealthEconomicsHackathon/inflatably

inflation_adjust_cost_custom <- function(from_year,
                                         to_year,
                                         from_cost,
                                         inflation_df) {

  if (from_year %% 1 != 0) stop("From date must be an integer valued whole year")
  if (to_year %% 1 != 0) stop("To date must be an integer valued whole year")
  if (from_cost < 0) stop("Cost must be non-negative")
  if(!from_year %in% inflation_df$year) stop("from year not in look-up table")
  if(!to_year %in% inflation_df$year) stop("to year not in look-up table")

  from_index <- inflation_df[inflation_df$year == from_year, "index"]
  to_index <- inflation_df[inflation_df$year == to_year, "index"]

  to_cost <- from_cost * (to_index/from_index)

  return(to_cost)
}


primary.cost <- data.table(mean_cost = c(100:120))


for (i in 1:nrow(primary.cost)){
  primary.cost[i, mean_i := inflation_adjust_cost_custom(2010,2018,primary.cost$mean_cost[i], pop.data)]
}
print(primary.cost)

```

## Other packages
'Rilostat' - bulk download of ilostat database. Useful for labour statistics.
https://cran.r-project.org/web/packages/Rilostat/Rilostat.pdf

## Other helpful resources
‘countrycode'- Convert Country Names and Country Codes into standardised formats.
https://cran.r-project.org/web/packages/countrycode/countrycode.pdf 